---
title: "ST 558 Project 1 XML"
author: "Ariel Macari, Xiyou Zhou"
output:
  html_document:
    cold_folding: hide
    toc: yes
    toc_float: FALSE
    toc_depth: 3
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning=FALSE, message = FALSE)
library(tidyverse)
library (XML)
library(DT)


```

# What is XML Data?

eXtensible Markup Language (XML) is a type of file used to store data.  The stored data includes annotations to format the data but remins invisible to the user end. XML Data can be opened in a web browswer or any text editor.  It is good use to store data for examples like a medical record.


[XML Data Example](https://www.ibm.com/support/knowledgecenter/en/SSLTBW_2.3.0/com.ibm.zos.v2r3.ib6p000/ib6p9c4m2.htm)

# Explanation on Packages
The following lists the packages that were used in this analysis

* XML - This package is a collection of functions 

* Tidyverse - The tidyverse is a system of packages that helps in read, clean, and present data.  The tidyverse includes the following packages     
    + ggplot2 - functions for creating graphics  
    + ddplyr - functions for data manipulation  
    + tidyr - functions to tidy data
    + readr- functions to read in files  
    + purr - functions to work with functions and vectors
    + tibble - a data frame with better options
    + stringr - functions to help working with strings
    + forcasts - functions to help with factors in R

* DT - This package allows R data ojbects to be displayed as tables on HTML pages.

# Data and Dataset
The dataset presents the age adjusted death rate for the 10 leading causes of deaths in the United States from 1999 - 2013.  The dataset includes the number of deaths by cause for each year and state.  Additionally the data set includes the Age-Adjusted Rate, which adjusts the the populations if each of the states had the same age distributions.


The data was found in [Dataset](https://catalog.data.gov/dataset/age-adjusted-death-rates-for-the-top-10-leading-causes-of-death-united-states-2013)

The data was read by the following:
* First, a reference to the website as an object called `health`. 
* Next, the xmL referenced was parsed inot a structure referencing the xml tree with `xmlParse`.
* Last the data was brought in a dataframe with the function `xmlToDataFrame`.  The function requests that a list of the xmL nodes to be processed be included.  The nodes resided in "//row//row"
* The following shows the code that was used to import the data into RStudio

``` {r Dataset, echo = TRUE}
health<-(file = "http://data.cdc.gov/api/views/bi63-dtpu/rows.xml?accessType=DOWNLOAD")
xmldoc <- xmlParse(health)
HealthData <- xmlToDataFrame(nodes = getNodeSet(xmldoc, "//row//row"))
(datatable(HealthData))
```


# Tables 
The function (include name here) calculates the % of overall deaths by the top 10 causes {need to creat this!}

```{r datatable}
knitr::kable(HealthData %>% filter(state =="United States" & cause_name=="All causes") %>% select(year, deaths, aadr) %>% arrange(year))


```


# Plots
at least 1 side by side bar
1 side by side box


```{r scatter}
Data<- HealthData %>% filter(state =="United States" & cause_name=="All causes") %>% select(year, deaths, aadr) %>% arrange(year)
ggplot(Data, aes(x=year, y = deaths, group=1))  + geom_line(lwd = 1.5, color = "red")

ggplot(Data, aes(x=year, y = aadr, group=1))  + geom_line(lwd = 1.5, color = "blue")
```