---
title: "ST 558 Project 1 XML"
author: "Ariel Macari, Xiyou Zhou"
output:
  html_document:
    cold_folding: hide
    toc: yes
    toc_float: FALSE
    toc_depth: 3
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning=FALSE, message = FALSE)
library(tidyverse)
library (XML)
library(DT)


```

# What is XML Data?

eXtensible Markup Language (XML) is a type of file used to store data.  The stored data includes annotations to format the data but remins invisible to the user end. XML Data can be opened in a web browswer or any text editor.  It is good use to store data for examples like a medical record.


[XML Data Example](https://www.ibm.com/support/knowledgecenter/en/SSLTBW_2.3.0/com.ibm.zos.v2r3.ib6p000/ib6p9c4m2.htm)

# Explanation on Packages
The following lists the packages that were used in this analysis

* XML - This package is a collection of functions 

* Tidyverse - The tidyverse is a system of packages that helps in read, clean, and present data.  The tidyverse includes the following packages     
    + ggplot2 - functions for creating graphics  
    + ddplyr - functions for data manipulation  
    + tidyr - functions to tidy data
    + readr- functions to read in files  
    + purr - functions to work with functions and vectors
    + tibble - a data frame with better options
    + stringr - functions to help working with strings
    + forcasts - functions to help with factors in R

* DT - This package allows R data ojbects to be displayed as tables on HTML pages.

# Data and Dataset
The dataset presents the age adjusted death rate for the 10 leading causes of deaths in the United States from 1999 - 2013.  The dataset includes the number of deaths by cause for each year and state.  Additionally the data set includes the Age-Adjusted Rate, which adjusts the the populations if each of the states had the same age distributions.


The data was found in [Dataset](https://catalog.data.gov/dataset/age-adjusted-death-rates-for-the-top-10-leading-causes-of-death-united-states-2013)

The data was read by the following:
* First, a reference to the website as an object called `health`. 
* Next, the xmL referenced was parsed inot a structure referencing the xml tree with `xmlParse`.
* Last the data was brought in a dataframe with the function `xmlToDataFrame`.  The function requests that a list of the xmL nodes to be processed be included.  The nodes resided in "//row//row"
* The following shows the code that was used to import the data into RStudio

``` {r Dataset, echo = TRUE}
health<-(file = "http://data.cdc.gov/api/views/bi63-dtpu/rows.xml?accessType=DOWNLOAD")
xmldoc <- xmlParse(health)
HealthData <- xmlToDataFrame(nodes = getNodeSet(xmldoc, "//row//row"))
HealthData$deaths<-as.numeric(HealthData$deaths)
HealthData$aadr<-as.numeric(HealthData$aadr)
(datatable(HealthData))
```


# Tables 


* The following shows the National number of deaths and aadr per years  
```{r datatable}
knitr::kable(HealthData %>% filter(state =="United States" & cause_name=="All causes"))
            
```

* A contingency table showing that the dataset includes 18 years of the top 10 causes of death by state is shown below.  The top 10 causes of death have remained the same throughout the years of 1999 - 2016.

```{r datatable1}
table(HealthData$state, HealthData$cause_name)
```

* The function (include name here) calculates the % of overall deaths by the top 10 causes {need to create this!}

```{r function}
H1d<-HealthData %>% filter(state =="United States") %>% group_by(cause_name) %>% arrange(cause_name,year) %>% select(cause_name, deaths,year) %>% group_by(year) %>% mutate(totdeaths=sum(deaths)) %>% mutate(Death_Percent = round(deaths/totdeaths*100,1) )
  
```

# Plots
at least 1 side by side bar, still working on this (actually using faet wrapping):

``` {r bar}
ggplot(H1d, aes(x=year ,y=Death_Percent)) + geom_bar(stat="identity") + facet_wrap(aes(cause_name)) 

```

1 side by side box


```{r scatter}
Data<- HealthData %>% filter(state =="United States" & cause_name=="All causes") %>% select(year, deaths, aadr) %>% arrange(year)
ggplot(Data, aes(x=year, y = deaths, group=1))  + geom_line(lwd = 1.5, color = "red")

ggplot(Data, aes(x=year, y = aadr, group=1))  + geom_line(lwd = 1.5, color = "blue")
```
