---
title: "ST 558 Project 1 XML"
author: "Ariel Macari, Xiyou Zhou"
output:
  html_document:
    cold_folding: hide
    toc: yes
    toc_float: FALSE
    toc_depth: 3
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning=FALSE, message = FALSE)
library(tidyverse)
library (XML)


```

# What is XML Data?

eXtensible Markup Language (XML) is a type of file used to store data.  The stored data includes annotations to format the data but remins invisible to the user end. XML Data can be opened in a web browswer or any text editor.  It is good use to store data for examples like a medical record.


[XML Data Example](https://www.ibm.com/support/knowledgecenter/en/SSLTBW_2.3.0/com.ibm.zos.v2r3.ib6p000/ib6p9c4m2.htm)

# Explanation on Packages
Create in list form

* XML - This package is a collection of functions 

* Tidyverse - The tidyverse is a system of packages that helps in read, clean, and present data.  The tidyverse includes the following packages     
    + ggplot2 - functions for creating graphics  
    + ddplyr - functions for data manipulation  
    + tidyr - functions to tidy data
    + readr- functions to read in files  
    + purr - functions to work with functions and vectors
    + tibble - a data frame with better options
    + stringr - functions to help working with strings
    + forcasts - functions to help with factors in R

# Dataset
The dataset presents the age adjusted death rate for the 10 leading causes of deaths in the United States from 1999 - 2013.  The dataset includes the number of deaths by cause for each year and state.  Additionally the data set includes the Age-Adjusted Rate, which adjusts the the populations if each of the states had the same age distributions.


The data was found in [Dataset](https://catalog.data.gov/dataset/age-adjusted-death-rates-for-the-top-10-leading-causes-of-death-united-states-2013)

The data was read by the following:
* First, a reference to the website as an object called `health`. 
* Next, the xmL referenced was parsed inot a structure referencing the xml tree with `xmlParse`.
* Last the data was convertd in a dataframe with the function `xmlToDataFrame`.  The function requests that a list of the xmL nodes to be processed be included.  The nodes resided in "//row//row"

``` {r Dataset, echo = TRUE}
health<-(file = "http://data.cdc.gov/api/views/bi63-dtpu/rows.xml?accessType=DOWNLOAD")
xmldoc <- xmlParse(health)
xmldf <- xmlToDataFrame(nodes = getNodeSet(xmldoc, "//row//row"))
```


# Data
Show the data (use datatable)
Create a new variable

# Tables 
show tables here

# Plots
at least 1 side by side bar
1 side by side box
scatter with coloring



